---
title: "simulation"
author: "Xiyue Zhang"
date: "2023-02-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stan Code
We pass the code as a string.
```{r}
stanmodelcode <-'
/* lg_t.stan */
functions {
// Vector square root
vector vecsqrt(vector x) {
	vector[dims(x)[1]] res;
	for (m in 1:dims(x)[1]){
		res[m] = sqrt(x[m]);
	}
return res; }
}
data {
  int<lower=0> N; 
  int<lower=0> J;    
  matrix[N,J] Z;
  vector[N] X; 
  vector[N] Y; 
}
parameters {
  real <lower=0> sigmax; 
  real <lower=0> sigmay;
  real <lower=0> sigmaalpha;
  real<lower=0> r1_global;
  real<lower=0> r2_global;
  real mualpha;
  real omegax;
  real omegay;
  real deltax;
  real deltay;
  real theta;
  vector[N] u; 
  vector[J] z;
  vector<lower=0>[J] r1_local;
  vector<lower=0>[J] r2_local;
  vector[J] alpha;
}
transformed parameters {
   real<lower=0> tau;
   vector<lower=0> [J] lambda;
   vector[J] beta;
   tau      = r1_global * sqrt(r2_global);
   lambda	  = r1_local .* vecsqrt(r2_local);
   beta	    =  z .* lambda*tau;
}
model {
  X 	~ normal(omegax+Z*alpha+u*deltax, sigmax);
  Y 	~ normal(omegay+Z*beta+X*theta+u*deltay, sigmay);
  u 	~ normal(0,1);
  
  for(k in 1:J){
    alpha[k] ~ normal(mualpha, sigmaalpha);
  }
// Constructing the prior for the lambda vector
    z ~ normal(0, 1);
    r1_local ~ normal(0.0, 1.0);
    r2_local ~ inv_gamma(0.5, 0.5);
// Constructing the prior for tau
    r1_global ~ normal(0.0, 1.0);
    r2_global ~ inv_gamma(0.5, 0.5);
    }
'
```
# generate function
The function is to generate the data based on our simulation rules.
```{r}
generate <- function(N,J,THETA,Z,deltamin,deltamax,deltaxsd,deltaysd,sigmaxmin,
            sigmaxmax,sigmaymin,sigmaymax,fracnonzeroAlpha,alphamean,alphasd,
            omegaxmean,omegaymean,omegaxsd,omegaysd,pleiotropy,pleiomin,
            pleiomax,fractionpleiotropic,pleiosd){
U 				= array(rnorm(N, mean=0, sd=1), dim = N)
MEANDELTA = runif(1, deltamin, deltamax)
DELTAX 		= rnorm(1, mean= MEANDELTA, sd=deltaxsd)
DELTAY 		= rnorm(1, mean= MEANDELTA, sd=deltaysd)
nonzeroAlpha  = round(fracnonzeroAlpha*J)
ALPHA = array(rep(0,J),dim=J)
ALPHA[sample(c(1:J),size=nonzeroAlpha,prob=rep(1,J),replace=F)] = 
          rnorm(nonzeroAlpha, mean= alphamean, sd=alphasd)
OMEGAX 		= rnorm(1,mean=omegaxmean,sd=omegaxsd)
OMEGAY 		= rnorm(1,mean=omegaymean,sd=omegaysd)
X 				= array(Z%*%ALPHA, dim = N)
X 				= X+OMEGAX+U*DELTAX
SIGMAX		= runif(1,sigmaxmin,sigmaxmax)
EPS 			= array(rnorm(N, mean=0, sd=SIGMAX), dim=N)
X 			  = X+EPS
MEANPLEIO = 0
PLEIO 		= runif(1,pleiomin,pleiomax)
if(pleiotropy == "POSITIVE")MEANPLEIO = PLEIO
if(pleiotropy == "NEGATIVE")MEANPLEIO = -PLEIO
BETA = array(rep(0,J), dim=J)
HOWMANYPLEIOTROPIC = round(fractionpleiotropic*J)
BETA[sample(c(1:J),size=HOWMANYPLEIOTROPIC,prob=rep(1,J),replace=F)] = 
          rnorm(HOWMANYPLEIOTROPIC,mean=MEANPLEIO,sd=pleiosd)
SIGMAY 		= runif(1,sigmaymin,sigmaymax)
EPSILON 	= array(rnorm(N, mean=0, sd=SIGMAY), dim=c(N,1))
esposizione 	= array(X*THETA, dim=c(N,1))
confo		= array(U*DELTAY, dim= c(N,1))
YBUF 		= OMEGAY+Z%*%BETA+esposizione+confo
YBUF 		= YBUF+EPSILON
Y 			= array(YBUF, dim=N)
dataset= list(X=X,Y=Y,Z=Z,U=U)
dataset
}
```
library loaded
```{r}
library(StanHeaders)
library(ggplot2)
library(rstan)
library(MendelianRandomization)
```

```{r}
f = function(X1)gsub("0\\.","\\.", X1)
```
# load matrix
The author did not upload the data, so we just generate a matrix with same dimension to test the code. 
Note: why a "random" matrix can still justify the result?
```{r}
Z<-matrix(nrow = 500,ncol=60)
n<-60
p<-runif(30000)
for (i in 1:500){
  for(j in 1:60){
    Z[i,j]<-rbinom(1,2,p[i])
  }
}
J = ncol(Z)
N = nrow(Z)
```
The average SNP-SNP correlation in this
particular set of data.
```{r}
mean(cor(Z))
```
# Experiment
use the experiment code provided
```{r}
THETA = 0.35
```


```{r}
simulated = generate(N,J,THETA,Z,
deltamin =-0.2,
deltamax =-0.1,
deltaxsd =0.02,
deltaysd =0.02,
sigmaxmin =0.05,
sigmaxmax = 0.15,
sigmaymin = 0.2,
sigmaymax = 0.4,
fracnonzeroAlpha = 0.3,
alphamean = 1.0, #-0.07
alphasd = 0.2,
omegaxmean = 3.3,
omegaymean = 0.9,
omegaxsd = 0.2,
omegaysd = 0.2,
pleiotropy = "POSITIVE",
pleiomin = -0.1,#0.006,
pleiomax = 0.3,#0.25, #0.36, #0.012,
fractionpleiotropic = 0.49,
pleiosd = 0.1#0.05
)
```

```{r}
X = array(simulated$X, dim = N)
Y = array(simulated$Y, dim = N)
Z = array(simulated$Z, dim = c(N,J))
U= array(simulated$U, dim = N)
```

```{r}
betaX <- array(NA, dim=J)
betaY <- array(NA, dim=J)
sebetaY <- array(NA, dim=J)
sebetaX <- array(NA, dim=J)
for(isnp in 1:J){
regX <- lm(X ~ Z[,isnp])
regY <- lm(Y ~ Z[,isnp])
betaX[isnp] <- summary(regX)$coefficients[2,1]
sebetaX[isnp] <- summary(regX)$coefficients[2,2]
betaY[isnp] <- summary(regY)$coefficients[2,1]
sebetaY[isnp] <- summary(regY)$coefficients[2,2]
}
```


```{r}
oggetto = mr_input(bx = as.numeric(betaX),
  bxse = as.numeric(sebetaX),
  by = as.numeric(betaY),
  byse = as.numeric(sebetaY),
  correlation = cor(Z),
  exposure = "X ", outcome = "Y",
          snps = colnames(Z))
```



```{r}
risultato = mr_allmethods(oggetto, method = "all")
thetamedianestimate = risultato$Values[2,2]
thetamedianestimate
```
```{r}
init_list = list(c1=list(theta=thetamedianestimate,
  beta=rep(0,J),alpha=betaX,deltax=0,
  deltay=0,u=rep(0,N)))
```

```{r}
dat = list(N=N, J=J,Z=Z,X=X,Y=Y)
```



# Stan 
```{r}
fit <- stan(model_code=stanmodelcode, data = dat,iter=1000,
  chains=1, init=init_list, verbose=F)
```
```{r}
theta = extract(fit,pars='theta',permuted=FALSE)
```


```{r}
motheta = monitor(theta)
```
```{r}
motheta
```

95\% confidence interval
```{r}
 cat("\n (", motheta[,4],",",
+     + motheta[,8],")",sep="")
```
Its posterior mean is:
```{r}
motheta[,6]
```



